---
title: "Week 9 Determinant and Sgn Functions"
output: pdf_document
---
Knitr file of the sgn function and determinant: \newline
Simple sgn function, returning 1 if an even number is passed and -1 if an odd is passed to it
```{r}
sgn <- function(n){
  return((-1)^(n))
}
```

Determinant function, first checks if the matrix has one column, in which case it returns the only entry (this seems easier than solving for a 2x2 at the end, etc.) as the determinant of a 1x1 matrix is that number. Then it finds the sum of all the determinant, utilizing the sgn function to decide whether they are +/-, multiplying the sgn times the current entry, and finally finding the determinant of whatever matrices are left. 
```{r}
determinant <- function(A){
  if(ncol(A)==1){
    return(A[1,1])
  }else{
    return(do.call(sum, lapply(1:ncol(A), function(i) sgn(i)*A[1,i] * determinant(A[-1,-i,drop=F]))))
  }
}
```

```{r}
noninvert <- matrix(c(1,2,3,4,5,6,7,8,9), byrow=T, ncol=3)
invert <- matrix(c(1,2,3,0,2,3,0,0,4), byrow=T, ncol=3)
```
A singular matrix:
```{r}
determinant(noninvert)
```
An uppert triangle matrix (to make demonstration easier):
```{r}
determinant(invert)
```