---
title: "Week 5 Assignment--IS 605"
author: "Brett Burk"
output: pdf_document
---
\noindent\textbf{1.a}
```{r}
A <- matrix(c(1,1,1,1,0,1,3,4), ncol=2)
b <- c(0,8,8,19)
```
Computing $A^{T}A$
```{r}
(AtA <- t(A) %*% A)
```
Computing $A^{T}b$
```{r}
(Atb <- t(A) %*% b)
```
\noindent\textbf{1.b}
Computing $(A^{T}A)^{-1} \cdot A^{T}b$ in order to solve for $\hat{x}$
```{r}
(solve(AtA) %*% Atb)
```
\noindent\textbf{1.c}
Computing the error $e$ given the vector $p$ in $A^{T}p$
```{r}
p <- c(1,5,13,17)
(Atp <- t(A) %*% p)
(x.hat <- solve(AtA) %*% Atp)
(e <- p - A %*% x.hat)
```
A simple floating point error keeps us from seeing all zeroes.\newline \newline
\noindent\textbf{1.d}
Calculating the error between $b$ and $p$
```{r}
(e = b - p)
```
\noindent\textbf{1.e}
Checking if e is orthogonal to p and each column of A
```{r}
all.equal(e %*% p, p %*% e)
all.equal(e %*% A[,1], A[,1] %*% e)
all.equal(e %*% A[,2], A[,2] %*% e)
```
\noindent\textbf{2}
```{r}
data <- read.table("auto-mpg.data")
names(data) <- c("displacement", "horsepower", "weight", "acceleration", "mpg")
A <- as.matrix(data[,1:4])
b <- as.matrix(data[,5])
AtA <- t(A) %*% A
Atb <- t(A) %*% b
(x.hat <- solve(AtA) %*% Atb)
```
Calculating the error of the prediction:
```{r}
(e <- sqrt(sum(((A %*% x.hat) - b)^2)))
```