---
output: pdf_document
---
\vspace*{\stretch{1.0}}
   \begin{center}
      \Large\textbf{Week 5 Assignment--IS 607}\\
      \large\textit{Brett Burk}
   \end{center}
   \vspace*{\stretch{2.0}}
\noindent 1. Questions:
i. How many total votes were there? \newline
ii. What percent voted yes? \newline
iii. Are the young more likely to vote yes? \newline
\newline
2.
```{r include=FALSE}
library(tidyr)
library(dplyr)
```

```{r}
messy <- data.frame(yes.no = c("yes", "no"), 
                    "ed.sixteen.twentyfour" = c(80100, 35900), 
                    "ed.twentyfive" = c(143000, 214800), 
                    "gl.sixteen.twentyfour" = c(99400, 43000), 
                    "gl.twentyfive" = c(150400, 207000))
messy
```
3.
```{r}
tidier <- messy %>%
  gather(group, votes, -yes.no) %>%
  arrange(yes.no)
tidier
```
4.i.
```{r}
total.votes <- select(tidier, votes) %>%
  sum()
total.votes
```
4.ii.
```{r}
percent.yes <-  filter(tidier, yes.no=="yes") %>%
  select(votes) %>%
  sum() %>%
  function(x) (x/total.votes) * 100
percent.yes
```
4.iii.
```{r}
# This could be vectorised, but for such a small sample, it seems pointless to not just copy and paste
young.votes <- filter(tidier, grepl("*sixteen.twentyfour", tidier$group))
young.total.votes <- select(young.votes, votes) %>%
  sum()
young.percent.yes <-  filter(young.votes, yes.no=="yes") %>%
  select(votes) %>%
  sum() %>%
  function(x) (x/young.total.votes) * 100
young.percent.yes

old.votes <- filter(tidier, grepl("*twentyfive", tidier$group))
old.total.votes <- select(old.votes, votes) %>%
  sum()
old.percent.yes <-  filter(old.votes, yes.no=="yes") %>%
  select(votes) %>%
  sum() %>%
  function(x) (x/old.total.votes) * 100
old.percent.yes
```
It seems that the young are more likely to prefer Cullen Skink \newline
\newline
5. This was a pretty efficient way to handle it, except for the last question, but I'm not sure how changing the table width/length would have much effected this (I love being able to pipe the data!). I love being able to write in plain english as well, it's much more natural using these packages.