---
title: "Best popular movies"
output: pdf_document
---


```{r, echo=FALSE}
library(ggplot2)
library(grid)
library(gridExtra)
library(plyr)
```

Here I loaded in the data (which is already built in)
```{r}
setwd("C:\\Users\\Brett\\Dropbox\\CUNY\\607")
m <- read.table("movies.tab", sep="\t", header=TRUE, quote="", comment="")
```
Popular movies can be easily defined by those that received the most votes. Thus it should be a function of rating and votes.

```{r}
popular <- ddply(m, "year", summarise, popularity=votes)
most.popular.year <- popular[which.max(popular$popularity),1]
most.popular.year
```
Best is less narrowly defined--likely factors can be rating--but we can also consider how many r10 ratings wer received (and the density around r10). 
Best can also be defined from the studios perspective by the budget being low, and the rating and votes being high (but for these examples, we'll ignore that).

Calculating by determining the popularity by multiplying the rating by the votes -- this is a simple measurement, as years that had lots of viewers will obviously be preferenced--even if they weren't as good:

```{r}
best1 <- ddply(m, "year", summarise, bestness=rating * votes)
ggplot(data=best1, aes(x=year, y=bestness)) + geom_point() + geom_point(aes(x=year[which.max(bestness)], y=bestness[which.max(bestness)]), col="Red")
best.year1 <- best1[which.max(best1$bestness),1]
best.year1
```

Calculating by using the mean of r10--movies which received the highest percentile of votes in the top 10%:

```{r}
best2 <- ddply(m, "year", summarise, bestness = mean(r10))
ggplot(data=best2, aes(x=year, y=bestness)) + geom_point() + geom_point(aes(x=year[which.max(bestness)], y=bestness[which.max(bestness)]), col="Red") + geom_smooth()
best.year2 <- best2[which.max(best2$bestness),1]
best.year2
```
Trying to find which year had the best most popular movies by multiplying the r10 (as a percent) times the amount of votes. This should tell us which movies had the most people giving it a rating in the top ten percent
```{r}
best.most.popular <- ddply(m, "year", summarise, best.pop = (mean(r10)/10) * votes)
ggplot(data=best.most.popular, aes(x=year, y=best.pop)) + geom_point() + geom_point(aes(x=year[which.max(best.pop)], y=best.pop[which.max(best.pop)]), col="Red") +           ggtitle("Best popular movies by year") +
ylab("R10 * Votes") +
  xlab("Year")
best.most.popular.year <- best.most.popular[which.max(best.most.popular$best.pop), 1]
best.most.popular.year
```

Thus we have determined that 2001 is the year that the best popular movies were made.