---
title: "week8RAssignment"
output: html_document
---

Step 1:

```{r}
flips <- rbinom(n=100, prob=0.5, size=1)
```

Step 2:
Assuming 0 is heads and 1 is tails

```{r}

sevenInARow = FALSE
for (i in 1:93){
  current <- flips[i:(i+6)]
  if(any(0 != current) == FALSE ||  any(1 != current) == FALSE){
    sevenInARow = TRUE
  }
}
(sevenInARow)
```

Step 3:
```{r}
flipCounter <- function(count){
  sevenInARowCount = 0
  for(x in 1:count){
    flips <- rbinom(n=100, prob=0.5, size=1)
    sevenInARow = FALSE
    for (i in 1:93){
      current <- flips[i:(i+6)]
      if(any(0 != current) == FALSE ||  any(1 != current) == FALSE){
        sevenInARow = TRUE
      }
    }
    if(sevenInARow == TRUE){
      sevenInARowCount = sevenInARowCount + 1
    }
  }
  return(sevenInARowCount/count)
}
```

Step 4:
```{r}
flipCounter(10)
flipCounter(100)
flipCounter(1000)
flipCounter(10000)

system.time(flipCounter(1000))
system.time(flipCounter(100000))
```
The estimate seems to slowly converge on 0.538. Fortunately, my machine has quite a bit of RAM (16gb) so it can handle R operations pretty well, 1000 loops executed in a few fractions of a second, but anything over that started to slow down. 100,000 took about 27 seconds to execute. (using system.time())

Using R's rle():
```{r}
fastFlipper <- function(count){
  sevenInARowCount <- 0
  for(x in 1:count){
    flips <- rbinom(n=100, prob=0.5, size=1)
    if (any(rle(flips)$lengths >= 7)){
      sevenInARowCount <- sevenInARowCount + 1
    }
  }
    return(sevenInARowCount/count)
}
fastFlipper(10)
fastFlipper(100)
fastFlipper(1000)
fastFlipper(10000)

system.time(fastFlipper(1000))
system.time(fastFlipper(100000))
```
Executing with the for loop, it took 27 seconds for 100,000, using rle, I can get the same results in less than 5 seconds.