Brett Burk
week 5 Assignment

1. Appending is find if you just want to put a few lines (or if you want to add lines to an existing variable [df, vector, etc.]). It should also be noted that appending can be slow due to the way R manages memory.
Data Frame Appending (which we can see is pretty laborious):
```{r}
egdfa <- data.frame(x = 0, y = 0, z = 0)
newRow <- c(0,0,0)
egdfa = rbind(egdfa, newRow)
egdfa = rbind(egdfa, newRow)
egdfa
```
Data Frame Preallocating:
```{r}
egdfp <- as.data.frame(matrix(0, nrow=3, ncol=3))
names(egdfp) <- c('x', 'y', 'z')
egdfp
```
Not only is this less cumbersome to type (especially for larger data sets) but it also executes much faster.
List Appending:
```{r}
eglista <- list(as.numeric("1"), "2", "3")
eglista <- c(eglista, "4")
eglista
```
List Preallocating:
```{r}
eglistp <- vector(mode = "list", length = 10)
eglistp
```
Vector Appending:
```{r}
egveca <- c("1", "2", "3")
egveca <- c(egveca, "4")
egveca
```
Vector Preallocating:
```{r}
egvecp <- vector(length = 10)
egvecp
```
2.
```{r}
library('MASS')
bostondf <- data.frame("Per Capita Crime" = Boston$crim, "Nitrogen Oxides PPM" = Boston$nox, "Property Tax per $10k" = Boston$tax)
head(bostondf)
```